---
title: "ST661: Assignment 3" 
author: "Francis Onyango Owino" 
date: "`r format(Sys.time(), '%d %B, %Y')`" 
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(GGally)
library(MASS)
library(mosaicData)
library(tidyverse)
```


#### Question 1

```{r}
head(Pima.tr, n = 2)
```

a.
```{r}
ggplot(data = Pima.tr,
       mapping = aes(x = bp, y = npreg)) +
  geom_point()
```

b.
```{r}
Pima.tr$npregf <- cut_interval(Pima.tr$npreg, n = 4,
                               labels = c("L1", "L2", "L3", "L4"))
Pima.tr |>
  head(n = 2)
```

c.
```{r}
ggplot(data = Pima.tr,
       mapping = aes(x = npregf, y = bp)) +
  geom_boxplot()
```

d.
```{r}
ggplot(data = Pima.tr,
       mapping = aes(x = glu, y = age, color = type)) +
  geom_point() +
  geom_smooth(method = lm)
```

e.
```{r}
ggplot(data = Pima.tr,
       mapping = aes(x = glu, y = age)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(vars(type))
```

#### Question 2

```{r}
data("Gestation")
Gestation |>
  head(n = 3)
```

a.
```{r}
ggplot(data = Gestation,
       mapping = aes(x = age, y = dage)) +
  geom_point() +
  labs(title = "Association between Mother's and Father's Age") +
  theme(plot.title = element_text(hjust = .5)) +
  geom_jitter()


ggplot(data = Gestation,
       mapping = aes(x = age, y = dage)) +
  geom_smooth(method = lm, se = F) +
  geom_jitter()

```

b.
```{r}
ggplot(data = Gestation,
       mapping = aes(race, fill = race))+
  geom_bar(position = "fill") +
  facet_wrap(vars(smoke)) +
  theme(legend.position = "none")
```

c.
```{r}
ggplot(data = Gestation,
       mapping = aes(x = race, y = age, fill = race)) +
  geom_boxplot()
```

```{r}
ggplot(data = Gestation,
       mapping = aes(race, age)) +
  geom_boxplot()
```

d.
```{r}
Gestation %>%
  filter(smoke %in% c("now", "never")) %>%
  ggplot(aes(x = race, y = wt, fill = smoke)) +
  geom_boxplot() +
  labs(fill = "Smoking Status") +
  theme_minimal()


```

#### Question 3
```{r}
hep <- read.csv("Hep2012.csv")
# names(hep)
```

a.
```{r}
points <- rowSums(hep[,-1], na.rm = T)
points
hep$Tot <- points
hep |> head()
names(hep)
```

b.
```{r}
hep |> 
  head(n = 2)
names <- hep |> 
  select(Athlete, Tot) |> 
  arrange(desc(Tot)) |> 
  head(n = 5) 

names <- names$Athlete
names <- trimws(names)
names
```

c.

```{r}
ggduo(data = hep,columnsY = "Tot",
      columnsX =  c("Long.Jump", "X100m.Hurdles", "X800m"))

```

d. 

```{r}
ggpairs(data = hep,
        columns = 2:8)
```
```{r}
athletes_with_zero_points <- hep[apply(hep[, -1], 1, function(x) any (x==0)), "Athlete"]

athletes_with_zero_points
trimws(athletes_with_zero_points[!is.na(athletes_with_zero_points)])
```

```{r}
hep[which.max(hep$Shot.Put), "Athlete"]

```

e.


```{r}
new_hep <- hep |> 
  select(X100m.Hurdles, High.Jump, Shot.Put, X200m, Long.Jump,Javelin, X800m)
ggparcoord(new_hep, groupColumn = "Athlete",
           scale = "globalminmax") + xlab("") + ylab("")
```
The use of globalminmax scale allows us to compare the patterns across different variables and makes the variables to be on a consitent scale for better visualization. 
Yes, athletes with zero points in any event can be seen from the plot. yes, It makes sense since i found some athlestes with zero scores in some events. 


f.

```{r}

hep <- hep |> arrange(desc(Tot))
hep$winners <- paste(1:nrow(hep), hep$Athlete, sep=" ")
hep$winners[-(1:3)] <- "Others"



hep <- hep |> arrange(Tot)

ggparcoord(hep,
           columns = 2:8,
           groupColumn = "winners",
           scale = "globalminmax") +
  xlab("") +
  ylab("") +
  scale_color_manual(values = c("red", "blue", "green", "gray"))

```




QUESTION FOUR


```{r}
water <- read_csv("BathingWaterQuality2013.csv")
water <- mutate(water,Date= lubridate::dmy(Date))
water4 <-
filter(water, Beach %in% c("Sutton", "Portmarnock" ,"Skerries" ,"Howth - Claremont"))
```

```{r}
attach(water4)
attach(water)
```
a.

```{r}
ggplot(data = water4,
       mapping = aes(x = Date, y = Enterococci, color = Beach)) +
  geom_point() +
  geom_line() +
  geom_hline(aes(yintercept = 100), color = "magenta")

```


b.
```{r}
ggplot(data = water4,
       mapping = aes(x = Date, y = E_Coli, color = Beach)) +
  geom_point() +
  geom_line() +
  geom_hline(aes(yintercept = 100), color = "magenta")

```

c.

```{r}
ggplot(data = water,
       mapping = aes(x = Date, y = Enterococci)) +
  geom_point(aes(col = Beach)) +
  geom_line() +
  facet_wrap(vars(Beach)) +
  geom_hline(yintercept = 100) 

```


d.

```{r}
ggplot(data = water, aes(x = Date)) +
  geom_point(aes(y = E_Coli, col = "E_Coli")) +
  geom_line(aes(y = E_Coli, col = "E_Coli")) +
  geom_point(aes(y = Enterococci, col = "Enterococci")) +
  geom_line(aes(y = Enterococci, col = "Enterococci")) +
  facet_wrap(~ Beach) +
  labs(title = "Water Quality over time",
       y = "Levels") +
  theme(plot.title = element_text(hjust = .5)) +
  geom_hline(yintercept = 100)

```

